---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# saveR

<!-- badges: start -->
<!-- badges: end -->

The goal of saveR is to provide the user with more flexibility during EDA by removing rows or columns from a data frame based on the percentage of data present in the row or column.

## Installation

You can install the development version of saveR from [GitHub](https://github.com/) with:

``` r
# install.packages("pak")
pak::pak("mkinlan/saveR")
```

## Example

Say you have a data set with nulls (aka, every data set) about products:

```{r example}
library(saveR)
## 
df <- data.frame(
  ID    = c(1234,2145,3468,2487,1222),
  Product_Name = c("ProdA", "ProdB", "ProdC", "ProdD", "ProdE"),
  Weight      = c(NA, 30, 22, 28, 35),
  Shipping_Zip  = c("12345", "23456", "34567", "45678", "56789"),
  Cost    = c(50000, NA, 55000, NA, NA),
  Phase    = c("Dev", "Prod", "Dev", "Obsolete", "Prod"),
  stringsAsFactors = FALSE  
)

df
```

You can easily clean a data set like this using a number of methods to remove nulls throughout the df:

```{r}
df1<-na.omit(df)
df1
```

or

```{r}
df2<-df %>% filter(is.na(df$Cost))
df2
```

```{r}
df3<-df %>% tidyr::drop_na()
df3
```

or

```{r}
#select rows that are complete for column "Income"
df4<-df[complete.cases(df[ , 5]),]
df4
```

But, maybe you don't want to get rid of all of the nulls in a row or column. For me, this is often because I'm exploring and cleaning a wide data set, and sometimes there so many columns that the easiest thing to do is to cut the whole thing down by chopping out the columns with blanks. 

But, by doing that, one risks the possibility of losing columns that are majority non-null, simply because they contain some null values. In situations like this, there might be some value in filtering out percentages of nulls. Then, the remaining data frame could be further assessed to determine if it would be worth it to take steps to try and fill in the remaining missing data. 

For example, in the data set above, Cost is missing a number of data points, but Weight is only missing one. Additionally, Weight likely refers to a product weight, which is probably standardized. It's likely that there is a table of product weights somewhere, and this could be used to fill in missing data.

Using the save_cols function from saveR would let us easily locate columns like Weight, where filling in missing data might be a fairly easy fix:

```{r}
result<-save_cols(df,0.75)
result
```

The resulting data frame has lost the Cost column, which was majority blank, but kept the Weight column, because 75% of the column contained non-null data. 

